---
title: "Text Mining Homework"
output:
  html_document:
    df_print: paged
author: "Lesley Duff"
date: "`r format(Sys.Date())`"
---
# Text Mining Homework

```{r}
library(janeaustenr)
library(tidytext)
library(tidyverse)
```

Using the dataset austen_books() from the package :

## 1. common words
Find the most common words in both Mansfield Park and Sense & Sensibility.

```{r}
#?janeaustenr
# https://github.com/juliasilge/janeaustenr
# mansfieldpark sensesensibility
#?austen_books

austen_books <- austen_books()
head(austen_books)
# columns text and book

austen_books() %>% 
    group_by(book) %>%
    summarise(total_lines = n())

mansfieldpark_book <- austen_books %>% 
  filter(book == "Mansfield Park")

mansfieldpark_common_words <- mansfieldpark_book %>% 
  unnest_tokens(word, text) %>% 
  count(word, book, sort = TRUE)
mansfieldpark_common_words

sensesensibility_book <- austen_books %>% 
  filter(book == "Sense & Sensibility")

sensesensibility_common_words <- sensesensibility_book %>% 
  unnest_tokens(word, text) %>% 
  count(word, book, sort = TRUE)
sensesensibility_common_words

```


## 2. common words no stop words
Find the most common words in both Mansfield Park and Sense & Sensibility, not including stop words.

## 3. sentiment words
Find the most common sentiment words in both Mansfield Park and Sense & Sensibility.

