---
title: "Video Game Sales"
output:
  html_document:
    df_print: paged
---
# Video Game Sales

```{r}
library(CodeClanData)
library(tidyverse)
```

```{r}
video_game_sales <- CodeClanData::game_sales

head(video_game_sales)
dim(video_game_sales)
```




```{r}
str(video_game_sales)
glimpse(video_game_sales)
```

```{r}
#video_game_sales %>% 
  #distinct(genre) 
  #distinct(publisher)
  #distinct(developer)
  #distinct(rating)
  #distinct(platform)
```

## Is there a relationship between user_score and critic_score?
```{r}
video_game_sales %>% 
 summarise(avg_user_score = mean(user_score),
            avg_critic_score = mean(critic_score))
  
video_game_sales %>% 
  ggplot(aes(x = user_score, y = critic_score)) +
  geom_point()
```

Yes there is a relationship. It looks like in general if `critic_score` is high 
then `user_score` will be high.

* Strongest when `user_score` > 5 and `critic_score` > 40
* Average `user_score` is 7.24
* Average `critic_score` is 74.28



## Do certain developers publish more of a certain genre of game?

```{r}
  
video_game_sales %>% 
#mutate(developer = as.factor(developer),
 #        genre = as.factor(genre)) %>% 
  ggplot(aes(x = developer, y=genre, fill = genre)) +
  geom_col(position = "dodge") #+
#  facet_grid(~developer ~ genre)
```
 Can see one with a lot of pink to be Sports but can't get which developer. Tried a lot of facet_wrap and facet_grid can't get something usable.
 
 Sorry, struggling with this spent 2 hours at this point(ugh!).
 
 Stopping :-(







