---
title: "Clustering"
author: "Lesley Duff"
date: "`r format(Sys.Date())`"
output:
  html_document:
    df_print: paged
---
# Clustering

## Task 1

Do some exploratory analysis on the data. Comment on findings.

```{r}
library(tidyverse)
```


```{r}
mall_customers <- read_csv("data/mall_customers.csv")

mall_customers <- mall_customers %>% 
  janitor::clean_names()
```


```{r}
head(mall_customers)

glimpse(mall_customers)

#skimr::skim(mall_customers) %>% View()

```
```{r}
mall_customers %>% 
  ggplot(aes(x = gender)) +
  geom_bar(colour = "white", fill="steelblue")

mall_customers %>% 
  ggplot(aes(x = as.integer(age))) +
  geom_histogram(binwidth = 5, colour = "white", fill="steelblue")

age_stats <- mall_customers %>% 
  summarise(min_age = min(age),
         max_age = max(age),
    mean_age = mean(age),
    annual_income_k = median(annual_income_k))
age_stats

mall_customers %>% 
  ggplot(aes(x = as.integer(annual_income_k))) +
  geom_histogram(binwidth = 5, colour = "white", fill="steelblue")+
  labs(
    x = "\nAnnual Income (£1000)",
    y = "Count\n"
  )

mall_customers %>% 
  ggplot(aes(x = spending_score_1_100)) +
  geom_histogram(binwidth = 5, colour = "white", fill="steelblue")+
  labs(
    x = "\nSpending Score (1-100)",
    y = "Count\n"
  )


```


No missing data.
Annual Income represents thousands of dollars
Spending Score is from 1 to 100.

From visualisation:
Females outnumber males in terms of visitors to the mall.
No children included Ages 18-70 with a mean of	38.85	

## Task 2

We are interested in creating a marketing campaign to target customers based on their spending score and annual income. Select an appropriate value for k and perform k-means clustering to find if there are meaningful clusters in the data
to target the customers.

```{r}
# spending score and annual income
mall_customers_spending_income <- mall_customers  %>% 
              select(spending = spending_score_1_100, 
                     income = annual_income_k)

head(mall_customers_spending_income)

mall_customers_spending_income %>% 
  ggplot(aes(x = spending, y = income)) +
  geom_point()
```
Visually looks like roughly 5 clusters.
```{r}
# Scaling
mall_customers_spending_income_scale <- mall_customers_spending_income %>%
                mutate_if(is.numeric, scale)

mall_customers_spending_income_scale


```




Task 3

Visualise the clustering for your chosen value of k.


Task 4

Do you think the clustering seems a good fit for this data?


Task 5

Comment on the attributes of one or two of the clusters (maybe even give them an appropriate label if you like - similar to section 4.1 of the ‘Segmentation & clustering intro’ lesson where we saw labels such as “Super Fans” and “Tag-Alongs”).