---
title: "Clustering"
author: "Lesley Duff"
date: "`r format(Sys.Date())`"
output:
  html_document:
    df_print: paged
---
# Clustering

## Task 1

Do some exploratory analysis on the data. Comment on findings.

```{r}
library(broom)
library(factoextra)
library(tidyverse)
```


```{r}
mall_customers <- read_csv("data/mall_customers.csv")

mall_customers <- mall_customers %>% 
  janitor::clean_names()
```


```{r}
head(mall_customers)

glimpse(mall_customers)

#skimr::skim(mall_customers) %>% View()

```
```{r}
mall_customers %>% 
  ggplot(aes(x = gender)) +
  geom_bar(colour = "white", fill="steelblue")

mall_customers %>% 
  ggplot(aes(x = as.integer(age))) +
  geom_histogram(binwidth = 5, colour = "white", fill="steelblue")

age_stats <- mall_customers %>% 
  summarise(min_age = min(age),
         max_age = max(age),
    mean_age = mean(age),
    annual_income_k = median(annual_income_k))
age_stats

mall_customers %>% 
  ggplot(aes(x = as.integer(annual_income_k))) +
  geom_histogram(binwidth = 5, colour = "white", fill="steelblue")+
  labs(
    x = "\nAnnual Income (£1000)",
    y = "Count\n"
  )

mall_customers %>% 
  ggplot(aes(x = spending_score_1_100)) +
  geom_histogram(binwidth = 5, colour = "white", fill="steelblue")+
  labs(
    x = "\nSpending Score (1-100)",
    y = "Count\n"
  )


```


No missing data.
Annual Income represents thousands of dollars
Spending Score is from 1 to 100.

From visualisation:
Females outnumber males in terms of visitors to the mall.
No children included Ages 18-70 with a mean of	38.85	

## Task 2

We are interested in creating a marketing campaign to target customers based on their spending score and annual income. Select an appropriate value for k and perform k-means clustering to find if there are meaningful clusters in the data
to target the customers.

```{r}
# spending score and annual income
mall_customers_spending_income <- mall_customers  %>% 
    column_to_rownames("customer_id") %>%  # for plotting

              select(spending = spending_score_1_100, 
                     income = annual_income_k)

head(mall_customers_spending_income)

mall_customers_spending_income %>% 
  ggplot(aes(x = spending, y = income)) +
  geom_point()
```
Visually looks like roughly 5 clusters.
```{r}
# Scaling
mall_customers_spending_income_scale <- mall_customers_spending_income %>%
                mutate_if(is.numeric, scale)

mall_customers_spending_income_scale


```
CHoose k

```{r}

# Set min & max number of clusters want to look at 
max_k <- 20 

k_clusters <- tibble(k = 1:max_k) %>%
  mutate(
    kclust = map(k, ~ kmeans(mall_customers_spending_income_scale, 
                             .x, nstart = 25)), 
    tidied = map(kclust, tidy),
    glanced = map(kclust, glance),
    augmented = map(kclust, augment, mall_customers_spending_income)
  )

k_clusters


clusterings <- k_clusters %>%
  unnest(glanced)

#clusterings

ggplot(clusterings, aes(x=k, y=tot.withinss)) +
  geom_point() +
    geom_line() +
    scale_x_continuous(breaks = seq(1, 20, by = 1))
```

Suggest 3 check 
```{r}
fviz_nbclust(mall_customers_spending_income_scale, 
             kmeans, 
             method = "wss", 
             nstart = 25)

fviz_nbclust(mall_customers_spending_income_scale, 
             kmeans, 
             method = "silhouette", 
             nstart = 25)

fviz_nbclust(mall_customers_spending_income_scale, 
             kmeans, 
             method = "gap_stat", 
             nstart = 25, 
             k.max = 20)
```

I'm choosing k=5


Task 3

Visualise the clustering for your chosen value of k.

```{r}
 clusterings %>% 
  unnest(cols = c(augmented)) %>%
  filter(k == 5) %>%
   ggplot(aes(x = spending, y = income, colour = .cluster, label = row.names(.))) +
  geom_point(aes(color = .cluster))+
  geom_text(hjust = -0, vjust = -1, size = 3)
```
## Task 4

Do you think the clustering seems a good fit for this data?

Yes, clear separation of clusters.

## Task 5

Comment on the attributes of one or two of the clusters (maybe even give them an appropriate label if you like - similar to section 4.1 of the ‘
Segmentation & clustering intro’ lesson where we saw labels such as 
“Super Fans” and “Tag-Alongs”).
I would need a better understanding of spending score. Need to go back to 
customer for clarification. I'm interpreting it as 
spending at this mall - higher meaning they purchase more at this particular mall.

3- Green - High income, Low spending
These could be incentivised to spend more as they can afford it.

2- olive - High income, High spending
1- red - Medium income, Medium spending
5- purple - Low income, Low spending
Less opportunities to increase sales

4 - blue - Low income, High spending