---
title: "Week 02, Day 01 Homework"
subtitle: "Credit card information and transactions"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
date: "`r format(Sys.Date())`"
author: "Lesley Duff"
---
# MVP

The data for the joining tasks is from [Kaggle](https://www.kaggle.com/ananta/credit-card-data) and contains synthetic (fake) credit card information and transactions. The data for the tidyr tasks is also synthetic.

```{r}
library(tidyverse)
library(janitor)
```
## Joins

### Question 1

Read in all 4 credit card transaction datasets and clean column names.

```{r}
cards <- read_csv("data/CardBase.csv")
customers <- read_csv("data/CustomerBase.csv")
fraud <- read_csv("data/FraudBase.csv")
transactions <- read_csv("data/TransactionBase.csv")
```

```{r}
cards <- cards %>%
  clean_names()

names(cards)

customers <- customers %>%
  clean_names()

names(customers)

# Not strictly speaking necessary for fraud right now but doing in case data 
# names are added to in future
fraud <- fraud %>%
  clean_names()

names(fraud)

transactions <- transactions %>%
  clean_names()

names(transactions)
```

### Question 2

Join the data containing card details and customer details by customer id, so 
that all records of card details and any matching records in customer details 
are kept. Before you run the code, think about how many rows you expect to see 
after joining.

```{r}
# cards: primary key: card_number, foreign key: cust_id
# customers: primary key: cust_id, foreign key: none
dim(cards)
dim(customers)

# Inner join will return all records in two tables that match on some condition.
# all records of card details and any matching records in customer details
# expecting rows = row count of cards

cards_customers <- inner_join(cards, customers, "cust_id")

dim(cards_customers) # rows here = rows in cards

cards_customers
```

### Question 3

Join the data containing fraud details with transaction details so all rows of 
both tables are kept. What does the resulting row number tell you?

```{r}
# fraud: primary key: transaction_id
# transactions: primary key
# Expecting rows to match rows of largest table

dim(fraud)
dim(transactions) # This table has more rows than fraud

# join fraud details with transaction details so all rows of both tables are
# kept
fraud_transactions <- full_join(fraud, transactions, "transaction_id")

dim(fraud_transactions) # this should equal transactions with most rows

fraud_transactions
```
Row number of `fraud_transactions` equals row number of the `transactions` 
table.

### Question 4 

Perform a filtering join between `cards` and `transactions` so that only the 
card details for cards that have made at least one transaction are preserved.

```{r}
dim(cards)
dim(transactions)

# ?semi_join
# returns rows from the cards table which have matching ones in the transactions
cards_transactions_has_transactions <- semi_join(cards, transactions,
  by = c("card_number" = "credit_card_id")
)

cards_transactions_has_transactions

dim(cards_transactions_has_transactions) # rows equals rows of cards

# just checking unique credit_card_id
# transactions %>%
#  distinct(credit_card_id)
#  row also equals rows of cards
```

## tidyr

### Question 5

Read in hat_observations and separate observation into two columns, `hat_colour` 
and `hat_type`.

```{r}
hat_observations <- read_csv("data/hat_observations.csv")

head(hat_observations)

# Add new columns of hat_colour and hat_type.
tidy_hat_observations <- hat_observations %>%
  separate(observation, c("hat_colour", "hat_type"), sep = ",")

head(tidy_hat_observations)
```

### Question 6

Unite `day`, `month`, and `year` columns into a column called `date` using a 
suitable separator. Then find the date where the most berets were observed.

```{r}
tidy_hat_observations_unite <- tidy_hat_observations %>%
  unite(date, c("year", "month", "day"), sep = "-") %>%
  mutate(date = as.Date(date)) %>%
  arrange(date)

head(tidy_hat_observations_unite)

tidy_hat_observations_unite %>%
  slice_max(observation_count, n = 1)
```

# Extension

## Joins

### Question 1

Can you join all 4 datasets together so that you’re left with a dataset that 
looks like below with 109 rows and 12 columns?

List of column names we have available:

  `cards`: "card_number"  "card_family"  "credit_limit" "cust_id"
  
  `customers`: "cust_id" "age" "customer_segment" "customer_vintage_group"
 
  `fraud`: "transaction_id" "fraud_flag"
  
  `transactions`: "transaction_id" "transaction_date" "credit_card_id"
    "transaction_value" "transaction_segment"

We want the following columns in this order:

* "card_number"  
* "card_family"  
* "credit_limit" 
* "cust_id" 
* "age" 
* "customer_segment" 
* "customer_vintage_group" 
* "transaction_id" 
* "transaction_date" 
* "transaction_value"
* "transaction_segment" 
* "fraud_flag"
  
This is combining all tables with   
`cards` + `customers` + `transactions` + `fraud`

```{r}
# inner_join on all foreign keys
cards_customers_transactions_fraud <- cards %>%
  inner_join(customers, cust_id, by = "cust_id") %>%
  inner_join(transactions, card_number,
    by = c("card_number" = "credit_card_id")
  ) %>%
  inner_join(fraud, transaction_id, by = "transaction_id")

cards_customers_transactions_fraud
```

## tidyr

### Question 2

Read in `exam_scores` and transform it into long format with two new columns 
`exam_question` and `score`. Then, using separate and select, remove superfluous
information from the values in `exam_question`

```{r}
exam_scores <- read_csv("data/exam_scores.csv")

spec(exam_scores)
head(exam_scores)

#  transform it into long format with two new columns exam_question and score
tidy_exam_scores <- exam_scores %>%
  pivot_longer(
    cols = c(
      "exam_Q1", "exam_Q2", "exam_Q3", "exam_Q4", "exam_Q5",
      "exam_Q6", "exam_Q7", "exam_Q8", "exam_Q9", "exam_Q10"
    ),
    names_to = "exam_question",
    values_to = "score"
  )

head(tidy_exam_scores)

# Then, using separate and select, remove superfluous information from the
# values in exam_question

# suggest removing prefix exam_Q to leave the digits
tidy_exam_scores <- tidy_exam_scores %>%
  separate(exam_question, c("exam", "question"), sep = "exam_Q") %>%
  
  # remove redundant exam column
  select(-exam) %>%
  
  # take the opportunity to convert question to numbers for any future functions
  # e.g. sorting by question number e.g. 1 and 10 would be next to each other
  # when it's a chr
  mutate(question = as.numeric(question)) %>%
  arrange(question)

tidy_exam_scores
```
