---
title: "Week 02, Day 02 Homework"
subtitle: "Tidyverse consolidation homework"
output:
  html_document:
    toc: true
    toc_float: true
#    number_sections: true
    df_print: paged
date: "`r format(Sys.Date())`"
author: "Lesley Duff"
---
# 1 Learning Objectives

Be able to explore and wrangle an unseen dataset so as to answer specific 
questions

# 2 Introduction

Social media is everywhere, and the ability to analyse data from it is 
invaluable. If you want to scrape data from Twitter yourself, you can use the 
rtweet package. All that is needed is a Twitter account, and you are good to go.

In this case, we have used the rtweet package to download 500 tweets from the 
@CodeClanScot twitter page.

In this homework, you will be working with data downloaded from the CodeClan 
Twitter page, to do some analysis of overall tweet metrics (likes and retweets),
as well as some analysis of the tweet and hashtag content. You’ll also be 
working with the strings contained within variables.
The overall goal of the homework is to practice everything you’ve learnt this 
week on an entirely new dataset.

# 3 MVP

## Question 1.

Load the code_clan_tweets.csv data. Find the number of rows, columns, and list 
all the variable names.

```{r}
library(tidyverse)
```

```{r}
code_clan_tweets <- read_csv("data/code_clan_tweets.csv")

dim(code_clan_tweets)
names(code_clan_tweets)
```

## Question 2.

Find the total number of favourites (stored in favorite_count) that CodeClan 
tweets have got. Don’t add a tweet’s favorite_count to the total when the tweet 
was quoted (see the is_quote variable).

```{r}
head(code_clan_tweets)
spec(code_clan_tweets)

# Find the total number of favourites (stored in favorite_count) that CodeClan 
# tweets have got.

code_clan_tweets %>%
  # Don’t add a tweet’s favorite_count to the total when the tweet was quoted 
  # (see the is_quote variable).
  select(favorite_count, is_quote) %>%
  filter(is_quote != TRUE) %>%
  summarise(total_number_favourites = sum(favorite_count))
```

## Question 3.

Summarise the mean number of retweets (retweet_count), split by the type of 
platform that was used to tweet (source). Again, omit quoted tweets from the 
mean.

```{r}
code_clan_tweets %>%
  group_by(source) %>%
  filter(is_quote != TRUE) %>%
  summarise(mean_number_retweets = mean(retweet_count)) %>%
  arrange(desc(mean_number_retweets))
```

## Question 4.

Count the total number of likes (i.e. total of favorite_count), split by media 
type, and arrange them from most likes to least. Assume that any tweets without 
a listed media type are of type “text”.

```{r}
code_clan_tweets %>%
  group_by(media_type) %>%
  # impute NAs into "text"
  # . ?replace
  mutate(media_type = replace(media_type, is.na(media_type), "text")) %>%
  summarise(total_number_likes = sum(favorite_count)) %>%
  arrange(desc(total_number_likes))
```

## Question 5.

Find the mean number of characters that a CodeClan tweet contains. You can 
either find a way to count the text in the text variable, or use the variable display_text_width which contains a count. Checking both would be a good idea if
possible to see if they match (but it’s not essential).

```{r}
code_clan_tweets %>%
  # ?str_length = length of string
  mutate(tweet_length = str_length(text)) %>%
  summarise(
    mean_number_characters = mean(tweet_length, na.rm = TRUE),
    mean_display_text_width = mean(display_text_width, na.rm = TRUE)
  )
```

## Question 6.

The `code_clan_info.csv` data file contains `status_url` along with other info. 
Load this in, and join it to the `code_clan_tweets` tibble, so that you have a 
`status_url` for each tweet. Decide which variable to join the tibbles on.

```{r}
code_clan_info <- read_csv(("data/code_clan_info.csv"))
```

```{r}
names(code_clan_tweets)

names(code_clan_info)

# both have a "tweet_id" column so will use that to join

# Checking primary keys for NA
code_clan_tweets %>%
  filter(is.na(tweet_id))

code_clan_info %>%
  filter(is.na(tweet_id))

# No NAs found proceed with mutating joins

# Extract just the status_url column
status_urls <- code_clan_info %>%
  select(tweet_id, status_url)
status_urls

# Combine tibble with status_url columns with code_clan_tweets
combined_tweets <- code_clan_tweets %>%
  left_join(status_urls, by = join_by(tweet_id))
combined_tweets
```


## Question 7.

From your new joined data, create a new tibble codeclan_hashtags containing only
tweet_id and hashtags, then convert hashtags into lowercase for analysis. Keep 
only those tweets with hashtags.

```{r}
codeclan_hashtags <- tibble(tweet_id = combined_tweets$tweet_id, hashtags = combined_tweets$hashtags)

tweets_with_hashtags <- codeclan_hashtags %>%
  filter(!is.na(hashtags)) %>%
  mutate(hashtags = str_to_lower(hashtags))

tweets_with_hashtags
```

# Extensions

## Question 8.

Some tweets have more than one hashtag, denoted by the `c(` symbols at the start 
of the string. Use the `str_sub()` and `str_detect()` functions and your 
`codeclan_hashtags` tibble from above to find all the cases in which the hashtag 
string begins with characters `c(`.
Hint
Try a regex pattern "c\\(". The double backslash is necessary in R to mark a 
backslash as itself being escaped.

```{r}
codeclan_hashtags %>%
  mutate(has_multiple_hashtags = str_detect(hashtags, "c\\(")) %>%
  # remove brackets
  mutate(hashtags = str_sub(hashtags, start = 3, end = -2)) %>%
  # all the cases in which the hashtag string begins with characters `c(`
  filter(has_multiple_hashtags == TRUE)
```


## Question 9.

Use the `str_detect()` function to find all tweets with text that mentions
“Edinburgh”, and count how many such tweets there are.
Hint
You may need to make sure your tweets are all the same case before searching.

```{r}
names(code_clan_tweets)
code_clan_tweets %>%
  #  ?str_detect
  # N.B. I dont' understand the hint if we make eg. everything lowercase then 
  # the mixed case "Edinburgh" wont match?

  mutate(has_edinburgh = str_detect(text, "Edinburgh")) %>%
  filter(has_edinburgh == TRUE) %>%
  summarise(count_tweets_have_edinburgh = n())
```

## Question 10.

Use str_extract_all() with a regular expression to find out which Twitter users 
CodeClan have been tweeting.
Hint
Think about all the different characters which could be in a Twitter username, 
do a bit of online research

```{r}
# Research: Help with username registration
# https://help.twitter.com/en/managing-your-account/twitter-username-rules
# Your username cannot be longer than 15 characters.
# A username can only contain alphanumeric characters (letters A-Z, numbers
# 0-9) with the exception of underscores, as noted above. Check to make sure
# your desired username doesn't contain any symbols, dashes, or spaces.

names(code_clan_tweets)
twitter_username_pattern <- "@([A-Za-z0-9_]+)"

code_clan_tweets %>%
  mutate(twitter_users = str_extract_all(text, twitter_username_pattern)) %>%
  unnest(twitter_users) %>%
  group_by(twitter_users) %>%
  # ?summarise
  summarise(twitter_users, n = n(), .groups = "keep") %>%
  distinct(n) %>%
  arrange(desc(n))
```
