---
title: "Sampling distributions and standard errors - mini lab"
subtitle: "NB Homework substitution instead of usual day 3"
output:
  html_document:
    df_print: paged
    toc: true
    toc_float: true
date: "`r format(Sys.Date())`"
author: "Lesley Duff"
---
# Sampling distributions and standard errors - mini lab


1. Load and explore the data held in `heart.csv`. A data dictionary is also available.  

```{r}
library(tidyverse)
```

```{r}
heart <- read_csv("data/heart.csv")

head(heart)
glimpse(heart)
```

  
2. Plot a histogram and/or boxplot for `thalach` (maximum heart rate achieved in
stress testing). How normal is the distribution?  
  
3. Also plot a barplot of `exang` (whether the patient displayed exercise 
induced angina)  
  
4. Now calculate population parameters: what is the mean `thalach` in the 
population. What proportion of patients have exercise induced angina (`exang`)?  
  
5. Now take a single sample of 50-patients from the population and calculate 
point estimates for mean `thalach` and proportion of `exang` from this sample 
(let's call these `mean_thalach` and `prop_exang` going forward).<br>
[**Hint** - use `rep_sample_n()` from the `infer` package to do this]  
  
6. What are the sampling errors of these point estimates?<br>[**Hint** - 
remember a *sampling error* is defined as the difference between a 
*point estimate* drawn from a sample, and a *population parameter*]  

7. Now we're going to repeatedly sample 50-patients from the population and 
build up sampling distributions for the `mean_thalach` and `prop_exang` point 
estimates. Take 5000 samples and store the sampling distributions in a variable
`rep_sample_50`.<br>[**Hint** - again, use `rep_sample_n()` to do this.]  
  
8. Visualise the sampling distributions of `mean_thalach` and `prop_exang` as 
histograms. How normal do the sampling distributions look?  

9. Obtain standard errors of the sampling distribution of `mean_thalach` and 
`prop_exang` directly as the standard deviations of the sampling distributions 
held in `rep_sample_50`  
  
10. Repeat steps 7. and 9. (omitting step 8.) once more with samples of size 
100-patients, and again with samples of size 200-patients. How do the standard 
errors of `mean_thalach` and `prop_exang` change as sample size increases? Does 
the *uncertainty* in the calculated statistics *increase* or *decrease* as 
sample size increases?<br>[**Hint** - it's OK to copy, paste and alter code to 
do this!]  





