---
title: "Sampling distributions and standard errors - mini lab"
subtitle: "NB Homework substitution instead of usual day 3"
output:
  html_document:
    df_print: paged
    toc: true
    toc_float: true
date: "`r format(Sys.Date())`"
author: "Lesley Duff"
---
# Sampling distributions and standard errors - mini lab


1. Load and explore the data held in `heart.csv`. A data dictionary is also available.  

```{r}
library(tidyverse)
```

```{r}
heart <- read_csv("data/heart.csv")

head(heart)
glimpse(heart)
str(heart)
```

  
2. Plot a histogram and/or boxplot for `thalach` (maximum heart rate achieved in
stress testing).
```{r}
heart %>% 
  ggplot(aes(x = thalach)) +
    geom_histogram(colour = "white", 
                   fill = "steelblue",
                   binwidth = 5) +
    labs(
      title = "Maximum heart rate achieved in stress testing",
      x = "\nMaximum heart rate",
      y = "Count\n"
    ) +
  theme_minimal()

heart %>% 
  ggplot(aes(x = thalach)) +
    geom_boxplot() +
    labs(
      title = "Maximum heart rate achieved in stress testing",
      x = "\nMaximum heart rate"#
    ) +
  theme_minimal()
```
 How normal is the distribution? 
 
```{r}
library(e1071)

heart %>%
  summarise(skewness = skewness(thalach, type = 1))
```
 
 This is a moderately left-skewed distribution.
  
3. Also plot a barplot of `exang` (whether the patient displayed exercise 
induced angina)  
```{r}
str(heart)
  heart %>% 
    mutate(exang = as.logical(exang)) %>% 
  ggplot(aes(x = exang, fill = exang)) +
    geom_bar(aes(fill = exang)) +
    labs(
      title = "Whether the patient displayed exercise induced angina",
      x = "\nPatient displayed exercise induced angina",
      fill = "Displayed angina"
    ) +
  theme_minimal()
```

4. Now calculate population parameters: what is the mean `thalach` in the 
population. What proportion of patients have exercise induced angina (`exang`)? 
```{r}
summary_popn <- heart %>%
  summarise(
    mean_thalach = mean(thalach), 
    prop_exang = sum(exang == 1) / n()
  )

summary_popn
```
* Mean `thalach` in the population: `r summary_popn$mean_thalach`.
* Proportion of patients with exercise induced angina (`exang`): `r summary_popn$prop_exang`.

  
5. Now take a single sample of 50-patients from the population and calculate 
point estimates for mean `thalach` and proportion of `exang` from this sample 
(let's call these `mean_thalach` and `prop_exang` going forward).<br>
[**Hint** - use `rep_sample_n()` from the `infer` package to do this]  
```{r}
library(infer)

sample_50 <- heart %>%
  # ?rep_sample_n
  rep_sample_n(size = 50, reps = 1) %>% 
#  get rid of error '-' only defined for equally-sized data frames
  ungroup()

sample_50

summary_sample_50 <- sample_50 %>%
  summarise(
    mean_thalach = mean(thalach), 
    prop_exang = sum(exang == 1) / n()
   ) 

summary_sample_50
```

  
6. What are the sampling errors of these point estimates?<br>[**Hint** - 
remember a *sampling error* is defined as the difference between a 
*point estimate* drawn from a sample, and a *population parameter*]  
```{r}
# Sampling error = point estimate - population parameter
sampling_error = summary_sample_50 - summary_popn
sampling_error
```

7. Now we're going to repeatedly sample 50-patients from the population and 
build up sampling distributions for the `mean_thalach` and `prop_exang` point 
estimates. Take 5000 samples and store the sampling distributions in a variable
`rep_sample_50`.<br>[**Hint** - again, use `rep_sample_n()` to do this.]  
```{r}
sample_50_5000 <- heart %>%
  rep_sample_n(size = 50, reps = 5000)

rep_sample_50 <- sample_50_5000 %>%
  summarise(
    mean_thalach = mean(thalach), 
    prop_exang = sum(exang == 1) / n()
   )

rep_sample_50
```

8. Visualise the sampling distributions of `mean_thalach` and `prop_exang` as 
histograms. How normal do the sampling distributions look?  
```{r}
rep_sample_50 %>% 
  ggplot(aes(x = mean_thalach)) +
    geom_histogram(colour = "white", 
                   fill = "steelblue") +
    labs(
      title = "Mean heart rate achieved in stress testing",
      x = "\nMean heart rate",
      y = "Count\n"
    ) +
  theme_minimal()
```

This is a unimodal distribution.

```{r}
rep_sample_50 %>% 
  ggplot(aes(x = prop_exang)) +
    geom_histogram(colour = "white", 
                   fill = "steelblue") +
    labs(
      title = "Proportion of `exang` heart rate achieved in stress testing",
      x = "\nProportion of `exang`",
      y = "Count\n"
    ) +
  theme_minimal()
```
This is a unimodal distribution.

9. Obtain standard errors of the sampling distribution of `mean_thalach` and 
`prop_exang` directly as the standard deviations of the sampling distributions 
held in `rep_sample_50`  
```{r}
std_errs <- rep_sample_50 %>%
  summarise(
    se_mean_thalach = sd(mean_thalach),
    se_prop_exang = sd(prop_exang)
  )

std_errs
```

10. Repeat steps 7. and 9. (omitting step 8.) once more with samples of size 
100-patients, and again with samples of size 200-patients. How do the standard 
errors of `mean_thalach` and `prop_exang` change as sample size increases? Does 
the *uncertainty* in the calculated statistics *increase* or *decrease* as 
sample size increases?<br>[**Hint** - it's OK to copy, paste and alter code to 
do this!]  

```{r}
sample_100_5000 <- heart %>%
  rep_sample_n(size = 100, reps = 5000)

rep_sample_100 <- sample_100_5000 %>%
  summarise(
    mean_thalach = mean(thalach), 
    prop_exang = sum(exang == 1) / n()
   )

rep_sample_100

std_errs_100 <- rep_sample_100 %>%
  summarise(
    se_mean_thalach = sd(mean_thalach),
    se_prop_exang = sd(prop_exang)
  )

std_errs_100

sample_200_5000 <- heart %>%
  rep_sample_n(size = 200, reps = 5000)

rep_sample_200 <- sample_200_5000 %>%
  summarise(
    mean_thalach = mean(thalach), 
    prop_exang = sum(exang == 1) / n()
   )

rep_sample_200

std_errs_200 <- rep_sample_200 %>%
  summarise(
    se_mean_thalach = sd(mean_thalach),
    se_prop_exang = sd(prop_exang)
  )

std_errs_200
```

std_errs_100:`r std_errs_100` vs std_errs_200:`r std_errs_200`.

Standard error roughly halves as the sample size doubles.

The smaller the standard error, the more representative the sample will be of the overall population. More certain = less uncertain.


