---
title: 'Decision trees'
output:
  html_document:
    df_print: paged
---
In this homework we will create a decision tree to see which factors are useful in predicting whether or not a passenger on the titanic will survive.

Run the code below before you begin:
```{r}
library(rpart)
library(rpart.plot)
library(tidyverse)

library(tidyverse)
titanic_set <- read_csv("data/titanic_decision_tree_data.csv")

shuffle_index <- sample(1:nrow(titanic_set))

# shuffle the data so class order isn't in order - need this for training/testing split later on
titanic_set <- titanic_set[shuffle_index, ]
```
```{r}
titanic_set %>%
  # skimr::skim() %>% View ()
  glimpse() %>%
  str()
```
## Question 1
Cleaning up the data is always the first step. 

```{r}
titanic_set %>%
  count(survived)

titanic_set %>%
  count(sex)

titanic_set %>%
  count(survived)

titanic_set %>%
  count(pclass)

titanic_set %>%
  count(age)

head(titanic_set)

dim(titanic_set)
titanic_clean <- titanic_set %>%
  # Take only observations which have a survived flag (i.e. that aren’t missing)
  filter(!is.na(survived)) %>%
  # Drop the NA
  filter(!is.na(age)) %>%
  filter(!is.na(embarked)) %>%
  filter(!is.na(fare)) %>%
  # Turn your important variables into factors
  # (sex, survived, pclass, embarkation)


  mutate(
    sex = factor(sex),
    #  survived_flag =  factor(recode(survived, "0" = "No", "1" = "Yes")),
    survived_flag = factor(survived, levels = c(0, 1), labels = c(
      "Died", "Survived"
    )),
    # did they survive, 0 = No, 1 = Yes

    class = factor(recode(pclass,
      "1" = "1st (Upper)",
      "2" = "2nd (Middle)",
      "3" = "3rd (Lower)"
    )),
    # class : Ticket class, 1 = 1st (Upper class),
    # 2 = 2nd (Middle Class), 3 = 3rd (Lower Class)

    port_embarkation = factor(recode(embarked,
      "C" = "Cherbourg",
      "Q" = "Queenstown",
      "S" = "Southampton"
    )),

    # port_embarkation: C = Cherbourg, Q = Queenstown, S = Southampton

    # Create an age_status variable which groups individuals under (and
    # including) 16 years of age into a category called “child” category
    # and those over 16 into a category called “adult”.
    age_status = if_else(age <= 16, "child", "adult")
  ) %>%
  select(
    -...1, -passenger_id, -name, -ticket, -fare, -cabin,
    -pclass, -survived, -embarked
    # Remove columns turned into factors
  )

head(titanic_clean)
# titanic_clean
```
## Question 2

```{r}
titanic_clean %>%
  ggplot(aes(x = sex, fill = survived_flag)) +
  geom_bar()

titanic_clean %>%
  ggplot(aes(x = age, fill = survived_flag)) +
  geom_histogram()

titanic_clean %>%
  ggplot(aes(x = sib_sp, fill = survived_flag)) +
  geom_histogram()

titanic_clean %>%
  ggplot(aes(x = parch, fill = survived_flag)) +
  geom_histogram()

titanic_clean %>%
  ggplot(aes(x = class, fill = survived_flag)) +
  geom_bar()

titanic_clean %>%
  ggplot(aes(x = port_embarkation, fill = survived_flag)) +
  geom_bar()

titanic_clean %>%
  ggplot(aes(x = age_status, fill = survived_flag)) +
  geom_bar()
```
Which variables do you think might be useful to predict whether or not people are going to die?

class (3rd class), sex (male), port_embarkation (Southampton), age (20-40),  
sib_sp (0), parch(0), age_status (adult)

## Question 3

Train/test split

Predicting whether will survive, (survived_flag = "Yes")

```{r}
# get how many rows we have in total to work out the percentage
n_data <- nrow(titanic_clean)

# create a test sample index
test_index <- sample(1:n_data, size = n_data * 0.2)

# create test set
titanic_test <- slice(titanic_clean, test_index)

# create training set
titanic_train <- slice(titanic_clean, -test_index)

titanic_test %>%
  janitor::tabyl(survived_flag)

titanic_train %>%
  janitor::tabyl(survived_flag)

# 0.1 Yes 25 0.3521127, Yes 263 0.4102964, 0.4102964-0.3521127  =  0.0581837
# 0.2 Yes 59 0.415493, Yes 229 0.4017544, 0.4017544-0.415493    = -0.0137386
# 0.3 Yes 91 0.42723, Yes 197 0.3947896, 0.3947896-0.42723.     = -0.0324404
# 0.4 Yes 121 0.4260563, Yes 167 0.3901869, 0.3901869-0.4260563 = -0.0358694
# 0.5 Yes 154 0.4325843, Yes 134 0.3764045,0.3764045-0.4325843  = -0.0561798

# 0.2 is smallest difference
```

## Question 4

Create your decision tree to try and predict survival probability using an 
appropriate method, and create a decision tree plot.

```{r}
titanic_fit <- rpart(
  formula = survived_flag ~ .,
  data = titanic_train,
  method = "class"
)

rpart.plot(titanic_fit,
  yesno = 2, # writes yes/no at all splits, 0 = no/yesno
  type = 2, # where conditions lie = on top
  fallen.leaves = TRUE, # leaves align at bottom (all in row)
  faclen = 10, # how many characters of categorical written
  digits = 2, # probability reported to x digits
  split.border.col = 3 # changes border colour
)
```



