---
title: "Probability homework"
subtitle: Week 06, Day 01
output:
  html_document:
    toc: true
    toc_float: false
    df_print: paged
date: "`r format(Sys.Date())`"
author: "Lesley Duff"
---
# Probability homework

MVP

## The Titanic

Some time after the disaster a journalist wrote that crew members were more 
likely to have survived the sinking than passengers. This was disputed by a 
descendant of one of the surviving crew members, who suggested that the opposite
was true, i.e. passengers were more likely to have survived than crew.

Calculate the conditional probabilities $p (survived | crew)$
 and $p (survived | passenger)$ to settle this dispute.

### Examine dataset

```{r}
# Load in libraries and helper functions ####
library(gtools)
library(tidyverse)
source('prob.R') 
```

```{r}
# Read titanic dataset ####
titanic <- read_csv("data/titanic.csv")
```

```{r}
# Examine titanic dataset variables in question ####

glimpse(titanic)

summary(titanic)

titanic %>% 
  group_by(survived) %>% 
  summarise(count = n())

titanic %>% 
  distinct(class) 
```
```{r}
# Create crew column ####
# Create new column to indicate whether each individual was a passenger (1st,
# 2nd or 3rd class) or a member of the crew (treat restaurant staff as crew)

titanic <- titanic %>% 
  mutate(
    crew = case_when(
      # passengers are not crew, Crew = FALSE
     class == "1st" | class == "2nd" | class == "3rd" ~ FALSE,
     # Crew = TRUE
     class == "engineering crew" ~ TRUE,
     class == "victualling crew" ~ TRUE,
     class == "restaurant staff" ~ TRUE,
     class == "deck crew" ~ TRUE,
     TRUE ~ FALSE), 
    .after = class
  )

#titanic

titanic %>% 
  group_by(crew) %>% 
  summarise(count = n()) %>% 
  ungroup()
```

### Calculate prob(survived ∣ crew)

$p(A | B) = \frac{p(A \cap B)}{p(B)}$, 
$p(survived | crew) = \frac{p(survived \cap crew)}{p(crew)}$

```{r}
prob_survived_given_crew = titanic %>%
  filter(crew == TRUE) %>%
  summarise(prob = sum(survived == "yes") / n())

prob_survived_given_crew
```

### Calculate prob(survived ∣ passenger)

$p(A | B) = \frac{p(A \cap B)}{p(B)}$, 
$p(survived | passenger) = \frac{p(survived \cap passenger)}{p(passenger)}$

```{r}
prob_survived_given_passenger = titanic %>%
  filter(crew == FALSE) %>%
  summarise(prob = sum(survived == "yes") / n())

prob_survived_given_passenger
```

### Titanic summary

| Person type | Probability type | Probability of surviving |
| ----------- | ---------------- | -----------------------: |
| Member of crew | $p(survived | crew)$  | `r prob_survived_given_crew$prob` |
| Passenger | $p(survived | passenger)$ | `r prob_survived_given_passenger$prob` |

Passengers had a 
`r (prob_survived_given_passenger$prob/prob_survived_given_crew$prob)` 
better chance of surviving than crew.

Therefore the descendant of one of the surviving crew members was correct and 
the journalist was wrong.

## Probability Consolidation Questions

### Question 1
If I roll a six-sided dice once, what is the probability of getting a number less than 5?

```{r}
# dice 6 faces 1-6
# = prob(event)=num(event outcomes)/num(all outcomes)
# p(< 5) 

# set up one roll of a dice
s_die <- rolldie(1)
s_die

num_less_than_five <- s_die %>%
  dplyr::filter(X1 < 5)

# calculate the probability
prob_num_less_than_five <- nrow(num_less_than_five) / nrow(s_die)
prob_num_less_than_five
```
Probability of getting a number less than 5: `r prob_num_less_than_five`

### Question 2

How many outcomes are there in the sample space of an experiment in which we 
toss three dice?

```{r}
# set up three roll of a dice
s_die <- rolldie(3)
s_die

num_outcomes_three_dice = nrow(s_die)
num_outcomes_three_dice
```
When we toss three dice there are `r num_outcomes_three_dice` possible outcomes.

### Question 3

In an experiment where we roll four dice:

i how many outcomes are in the event where the total of the four dice equals 5
, 6 or 7)?

Hint
For part i, rowSums(.) may help you setup up the necessary random variable 
total (or you can use pivot_longer())

```{r}
# set up four roll of a dice
s_die <- rolldie(4)

total_four_dice <- s_die %>% 
  mutate(dice_total = rowSums(.)) %>% 
  filter(dice_total %in% 5:7)

outcomes_four_dice_five_to_seven = nrow(total_four_dice)
```

Where the total of the four dice equals 5, 6 or 7 there are 
`r outcomes_four_dice_five_to_seven` outcomes.

ii what is the probability of this event?
```{r}
# p(event)=n(event outcomes)/n(all outcomes)
# p(four dice equals 5, 6 or 7 )
outcomes_four_dice = nrow(s_die)
prob_four_dice_five_to_seven = outcomes_four_dice_five_to_seven/outcomes_four_dice
prob_four_dice_five_to_seven
```
Probability of four dice equals 5, 6 or 7 = `r prob_four_dice_five_to_seven`.

We assume equally likely outcomes of a dice throw.

### Question 4

Five chairs are placed in a row in a room, along with five people of differing 
height. The people sit down randomly on the chairs. What is the probability that they sit in order of decreasing height, left to right?

Hint
This is an arrangement where order matters. Of all the possible arrangements, 
how many are correct?

```{r}
# Five chairs
# Five people
# Order matters = permutation

# 1 2 3 4 5 
#p(highest height)= 1/5, 1/4, 1/3, 1/2, 1

library(gtools)

people <- c("Alice (1)", "Bob(2)", "Carole(3)", "David(4)", "Elaine(5)")
# Choose one at a time, r = 1
# Don't replace an employee once chosen: repeats.allowed = FALSE
seated_people <- as_tibble(permutations(n = length(people), 
                                  r = 5, 
                                  v = people, 
                                  repeats.allowed = FALSE))
seated_people

# how many permutations do we have?
total_permutations = nrow(seated_people)
total_permutations

# probability that they sit in order of decreasing height
# {(5,4,3,2,1),... }
# the first one is in order of decreasing height
# p(event)=n(event outcomes)/n(all outcomes)
# p(order of decreasing height, left to right)
prob_decreasing_height_left_to_right = 1 / total_permutations
```
Probability that they sit in order of decreasing height, left to right: 
`r prob_decreasing_height_left_to_right`
