---
title: "NYC Citi Bike trips"
subtitle: "Communicating Data"
output:
  html_document:
    toc: true
    toc_float: true
    df_print: paged
date: "`r format(Sys.Date())`"
author: "Lesley Duff"
---
# Introduction

[*Citi Bike*](https://ride.citibikenyc.com/about) is the largest [bike share system](https://en.wikipedia.org/wiki/Bicycle-sharing_system) in the USA and consists of a fleet of bikes available to users at docking stations throughout New York City. Bikes can be returned to any other station. Citi Bike is available for use 24 hours/day, 7 days/week, 365 days/year.

This report uses a sample dataset from NYC Citi Bike usage of 10 bikes throughout 2018. The data includes event data on each trip, including the trip's start and end times and locations. 

```{r}
library(skimr)
library(tidyverse)
library(tsibbledata)
```

```{r}
# ?nyc_bikes comes from the tsibbledata package
nyc_bikes_df <- nyc_bikes

head(nyc_bikes_df)
glimpse(nyc_bikes_df)

skim(nyc_bikes_df)

# We have 4268 rows, 12 columns
# We have no NAs 
# Datetime columns: start_time, stop_time
# birth_year is numeric

```

## The pattern of bike hires

over time (e.g. within a year, month, week, or day)?

## bike rider demographics

Do bike hire patterns differ between bike rider demographics? (e.g. gender, type of trip, age)
Any other insights?