---
title: "Advanced `ggplot2` homework"
output:
  html_document:
    df_print: paged
---
# MVP





**Question 1.**  

Load in the libraries you need, and look at the `backpack` data. Take note of variable names and dimensions. 

```{r}
library(CodeClanData)
library(tidyverse)
```

```{r}
head(backpack)
str(backpack)

# Variable names
names(backpack)

# Dimensions
dim(backpack)

#backpack %>% 
#  distinct(major)# %>% # Philosophy/GRC/Bio
 # distinct(sex) # Female/Male
 # distinct(status) # U/G
 # distinct(year). # 0-6
#distinct(units) #0 -19
```

**Question 2.**

What is the distribution of backpack weights that students are carrying? Choose an appropriate graph to show the distribution. 


**Question 3.**  

Are there differences in backpack weight between sex?   


**Question 4.**  

What is the relationship between body weight and backpack weight?


**Question 5.**  

Is this relationship different between men and women? Choose an appropriate plot and interpret your findings.  


**Question 6.**  

What **proportion** of students report back pain?   


**Question 7.**  

Is this proportion different between men and women? 


**Question 8.**  

Now you've investigated your data separately, create plots to describe the relationship between backpack weight, body weight, gender and reported backpain.

Try both `backpack_weight` and `body_weight` as well as the `ratio`.   

*Hint: faceting would be useful here*.  



