---
title: "Advanced `ggplot2` homework"
output:
  html_document:
    df_print: paged
---
# MVP

**Question 1.**  

Load in the libraries you need, and look at the `backpack` data. Take note of variable names and dimensions. 

```{r}
library(CodeClanData)
library(tidyverse)
```

```{r}
head(backpack)
str(backpack)

# Variable names
names(backpack)

# Dimensions
dim(backpack)

#backpack %>% 
#  distinct(major)# %>% # Philosophy/GRC/Bio
 # distinct(sex) # Female/Male
 # distinct(status) # U/G
 # distinct(year). # 0-6
#distinct(units) #0 -19
```

**Question 2.**

What is the distribution of backpack weights that students are carrying? Choose an appropriate graph to show the distribution. 

```{r}
#View(backpack)
backpack %>% 
  summarise(avg_backpack_weight = mean(backpack_weight),
            median_backpack_weight = median(backpack_weight),
            min_backpack_weight = min(backpack_weight),
            max_backpack_weight = max(backpack_weight))

backpack %>% 
  ggplot(aes(x = backpack_weight)) +
 # ?geom_histogram
  geom_histogram(binwidth = 5, col = "white", boundary = 5) +
#  geom_histogram(binwidth = 5, col = "white") +
  theme_minimal()+
  labs(x = "\nBackpack Weight (lbs)",
       y = "Count\n",
       title = "Backpack Weight by Five Pound Intervals")
```

* Most commonly backpacks will weigh greater than 5 pounds and less than 15 pounds. 
* The lightest is 2 pounds
* The heaviest is 35 pounds
It seems that this dataset is North American in origin as the likely units of weight are in pounds. [Backpack](https://en.wikipedia.org/wiki/Backpack)

**Question 3.**  

Are there differences in backpack weight between sex?   

```{r}
backpack %>% 
  select(sex, backpack_weight) %>% 
  group_by(sex) %>% 
  summarise(count = n(),
            avg_backpack_weight = mean(backpack_weight)) %>% 
  ungroup()
  
backpack %>% 
  ggplot(aes(x = backpack_weight, fill = sex)) +
 # ?geom_histogram
  geom_histogram(binwidth = 5, col = "white", boundary = 5) +
    facet_wrap(~sex)+
#  geom_histogram(binwidth = 5, col = "white") +
  theme_minimal()+
  labs(x = "\nBackpack Weight (lbs)",
       y = "Count\n",
       title = "Backpack Weight by Five Pound Intervals and Sex")
```
* There are more females (55) than males (45). 
* Average backpack weight for females is 11.04 lbs.
* Average backpack weight for males is 12.42 lbs.
* More males carry backpacks in the 15-20 lb range than females.
* More females carry backpacks in the 20-25 lb range than males. 
* Most common backpack weight for females is> 5 lbs < 10 lbs.
* Most common backpack weight for males is > 10 lbs < 15 lbs.



**Question 4.**  

What is the relationship between body weight and backpack weight?


**Question 5.**  

Is this relationship different between men and women? Choose an appropriate plot and interpret your findings.  


**Question 6.**  

What **proportion** of students report back pain?   


**Question 7.**  

Is this proportion different between men and women? 


**Question 8.**  

Now you've investigated your data separately, create plots to describe the relationship between backpack weight, body weight, gender and reported backpain.

Try both `backpack_weight` and `body_weight` as well as the `ratio`.   

*Hint: faceting would be useful here*.  



