---
title: "Homework Week 05, Day 02"
output:
  html_document:
    df_print: paged
---
# Advanced ggplot2 homework

You’ll be looking at the backpack dataset contained within the CodeClanData package using tidyverse. You’ll need to load the appropriate packages first.In all plots you make, ensure your labels are appropriate


## Question 1.

Load in the libraries you need, and look at the backpack data. Take note of variable names and dimensions.

```{r}
library(CodeClanData)
library(tidyverse)
```
```{r}
head(backpack)
str(backpack)

dim(backpack) # 100 rows 9 columns
backpack
```

## Question 2.

What is the distribution of backpack weights that students are carrying? Choose an appropriate graph to show the distribution.

```{r}
backpack
backpack_weight

# histogram, linel scatter plot (if 2 variables)

ggplot(backpack) +
  aes(x = branch, y = sales) +
  geom_col() +
  coord_flip()
```


## Question 3.

Are there differences in backpack weight between sex?

```{r}

```

## Question 4.

What is the relationship between body weight and backpack weight?

```{r}

```

## Question 5.

Is this relationship different between men and women? Choose an appropriate plot and interpret your findings.

```{r}

```

## Question 6.

What proportion of students report back pain?

```{r}

```

## Question 7.

Is this proportion different between men and women?

```{r}

```

## Question 8.

Now you’ve investigated your data separately, create plots to describe the relationship between backpack weight, body weight, gender and reported backpain.

Try both backpack_weight and body_weight as well as the ratio.

Hint: faceting would be useful here.

```{r}

```

