---
title: "Features and elements of multiple regression"
output:
  html_document:
    df_print: paged
date: "`r format(Sys.Date())`"
author: "Lesley Duff"
---
# Glossary of Terms
- *Regression*: the process of finding quantifiable relationships between "variables", i.e. saying that we can approximate a (continuous numeric) variable by a weighted combination of other variables
- *Linear Regression*: a formula with additive weights of predictor variables
- *Explanatory/predictor/independent variables*: the variables (columns, measurements of real-world phenomena) we use to try and estimate the thing we're interested in...
- *Outcome/response/dependet variablE (singular)*: the thing we're trying to estimate
- *Multiple (Linear) Regression*: finding a regression using more than one explanatory/predictor/independent variables

===
# Performing regression step-by-step
## Familiarise ourselves with the data and how it relates to what we know about the real-world process

=========
# Multiple regression

We have data for [California Housing Prices](https://www.kaggle.com/datasets/camnugent/california-housing-prices).

"Median house prices for California districts derived from the 1990 census."

The data contains information from the 1990 California census and pertains to the houses found in a given California district and some summary stats about them based on the 1990 census data. Columns: 

* longitude
* latitude
* housing_median_age
* total_rooms
* total_bedrooms
* population
* households
* median_income
* median_house_value
* ocean_proximity

```{r}
library(GGally)
library(tidyverse)
```

```{r}
# Read clean data
housing_prices <- read_csv("clean_data/housing_prices.csv")
```
## total_rooms of houses and total_bedrooms correlation?

```{r}
# We expect the total_rooms of houses to be strongly correlated with total_bedrooms. Use ggpairs() to investigate correlations between these two variables.


ggpairs(housing_prices)






# Check numerical correlations
housing_prices %>%
  summarise(correlation = cor(total_rooms, total_bedrooms))

# 0.9307124	 = 	very strong positive correlation
```


```{r}
ggpairs(). #?ggpairs
```

